### 2016-08-16 latex/input/related.tex
paper draft v1.1: most minor edits completed

The Multiple Instance Learning (MIL) problem was first described by Dietterich in the context of drug activity prediction \cite{dietterich97}. In this problem, we want to figure out which of a number of different molecules will bind to a target "binding site". Each molecule can assume a number of different 3-dimensional shapes, so even for molecules that are known to bind to the binding site, it is not necessarily known which formation of the molecule succeeds in binding. If even one shape of a given molecule binds to the binding site, it is considered a "good" molecule \cite{dietterich97}. Thus, in the original formation of MIL, a bag is classified as positive if one or more instances within it is positive, while a negative bag contains only negative instances. This is commonly referred to as the "standard multiple instance assumption" \cite{amores13}. In the original paper, Dietterich develops a solution based on axis-parallel rectangles to solve this problem, and the MIL approach was shown to be significantly more effective than a standard supervised learning approach \cite{dietterich97}. In the late 1990s and early 2000s, a number of different approaches were developed for the original MIL problem, such as Diverse Density (DD) \cite{perez98}, EM-DD \cite{zhang01}, MI-SVM \cite{andrews02}, sbMIL \cite{bunescu07}, and MILES \cite{wang06}. A recent review of MIL by Amores created a taxonomy of these various methods and compared their effectiveness for classification \cite{amores13}. Most of these classic methods follow the standard assumption, which is harmful in domains in which negative bags may contain some positive instances.
%

More recently, there has been increased interest in different formulations of the MIL problem. For instance, the problem of "key instance detection" \cite{zhou12} revolves around finding the instances that contribute the most to bag labels. A recent study focused on a formulation of the MIL problem in which bags with negative labels can contain some "positive" instances, and developed a general cost function for determining individual instance labels from group labels \cite{kotzias15}. This is significant in metagenomics because, while some diseases are caused by a single pathogen, many arise from a combination of many factors, and even patients that are "healthy" may contain small amounts of pathogens that are normally associated with disease. In contrast with the standard assumption, this can be referred to as the "collective" assumption \cite{amores13}. Additionally, it is helpful to discover which microbes and which functional attributes of those microbes lead to disease, making instance level information significant in this domain.
%

While some of the above methods follow the standard assumption and others follow the collective assumption, they all treat bag labels simply as aggregations of instance labels and thus focus only on comparisons between individual instances and not entire bags or groups of instances. For methods following the standard assumption, the aggregation function is simply an OR function: if any of the instances in a bag are positive, the entire bag is positive. For methods following the collective assumption, the aggregation function is often based on an averaging of instance labels. Regardless of the problem assumption, methods that rely only on comparisons between individual instances are referred to as "Instance Space" by Amores, and he showed that this paradigm was generally significantly less effective than the two other paradigms, "Bag Space" and "Embedded Space" \cite{amores13}. Bag Space methods define a distance or kernel function that determine the similarity between bags, while Embedded Space methods map bags into feature vectors which can then be used for classifiers \cite{amores13}. Embedded Space methods can be further divided into two subcategories: methods that simply aggregate information about all instances in a bag without differentiating them, and "Vocabulary-based" methods that group certain similar instances together and then use those groups to form the feature vector \cite{amores13}. We use a vocabulary-based method in this paper, because having information about groups of similar sequence reads can be biologically important, as explained further in the clustering section.

@@

Some of the popular genome and metagenome assembly approaches include SOAPdenovo2 \cite{luo12}, IDBA-UD \cite{peng12}, Velvet \cite{zerbino08}, MetaVelvet \cite{namiki12}, ABySS \cite{simpson09}, and Ray Meta \cite{boisvert12}. We used SOAPdenovo2 in our study, because it was the assembler used in the original MGWAS study \cite{qin041012} and because it has been shown to be one of the fastest assembly algorithms \cite{peng12}. SOAPdenovo2 first constructs a type of directed graph called a \emph{de Bruijn} graph that represents the overlaps between different sequences \cite{li10}. Reads are divided into strings of length K called \emph{k-mers}; these k-mers are the nodes of the graph \cite{zerbino08}. The choice of K is up to the user, and is important for having good assembly results. The k-mer nodes in the graph have an edge between them if a read contains those k-mers in order with an overlap of K-1 nucleotides, and the direction of the edge indicates in which order the k-mers appear \cite{zerbino08}. SOAPdenovo2 then cleans this graph by removing nodes/sequences with few or no connections with other sequences, eliminating "tips" that represent likely sequencing machine errors, and removing redundant edges \cite{li10}. The contigs are then formed by combining reads according to the de Bruijn graph: each contig represents a directed path in the graph \cite{zerbino08}. There are variations on how exactly de Bruijn graphs are represented and implemented, but Figure \ref{de-bruijn} illustrates the general process.

@@

UCLUST \cite{Edgar10}, CD-HIT \cite{Li01072006}, mothur \cite{schloss2009introducing}, DOTUR \cite{schloss2005introducing}, CROP \cite{Hao01032011}, MC-MinH \cite{sdm2013a}, and 
MC-LSH \cite{bibm2012} are some of the popular  nucleotide sequence clustering approaches. We use UCLUST within our study, which is one of the most widely used and cited metagenome clustering methods and has been shown to be amongst the most effective in terms of speed and accuracy in benchmarking studies \cite{bonder090112, sun042711}. 
%

>>>>

The Multiple Instance Learning (MIL) problem was first described by Dietterich in the context of drug activity prediction \cite{dietterich97}. In this problem, we want to figure out which of a number of different molecules will bind to a target ``binding site". Each molecule can assume a number of different 3-dimensional shapes, so even for molecules that are known to bind to the binding site, it is not necessarily known which formation of the molecule succeeds in binding. If even one shape of a given molecule binds to the binding site, it is considered a ``good" molecule \cite{dietterich97}. Thus, in the original formation of MIL, a bag is classified as positive if one or more instances within it is positive, while a negative bag contains only negative instances. This is commonly referred to as the ``standard multiple instance assumption" \cite{amores13}. In the original paper, Dietterich develops a solution based on axis-parallel rectangles to solve this problem, and the MIL approach was shown to be significantly more effective than a standard supervised learning approach \cite{dietterich97}. In the late 1990s and early 2000s, a number of different approaches were developed for the original MIL problem, such as Diverse Density (DD) \cite{perez98}, EM-DD \cite{zhang01}, MI-SVM \cite{andrews02}, sbMIL \cite{bunescu07}, and MILES \cite{wang06}. A recent review of MIL by Amores created a taxonomy of these various methods and compared their effectiveness for classification \cite{amores13}. Most of these classic methods follow the standard assumption, which is not useful in real domains in which negative bags may contain some positive instances.
%

More recently, there has been increased interest in different formulations of the MIL problem. For instance, the problem of ``key instance detection" \cite{zhou12} revolves around finding the instances that contribute the most to bag labels. A recent study focused on a formulation of the MIL problem in which bags with negative labels can contain some positive instances, and developed a general cost function for determining individual instance labels from group labels \cite{kotzias15}. This is significant in metagenomics because, while some diseases are caused by a single pathogen, many arise from a combination of many factors, and even patients that are healthy may contain small amounts of pathogens that are normally associated with disease. In contrast with the standard assumption, this can be referred to as the ``collective" assumption \cite{amores13}. Additionally, it is helpful to discover which microbes and which functional attributes of those microbes lead to disease, making instance level information significant in this domain.
%

While some of the above methods follow the standard assumption and others follow the collective assumption, they all treat bag labels simply as aggregations of instance labels and thus focus only on comparisons between individual instances and not entire bags or groups of instances. For methods following the standard assumption, the aggregation function is simply an OR function: if any of the instances in a bag are positive, the entire bag is positive. For methods following the collective assumption, the aggregation function is often based on an averaging of instance labels. Regardless of the problem assumption, methods that rely only on comparisons between individual instances are referred to as ``Instance Space" by Amores, and he showed that this paradigm was generally significantly less effective than the two other paradigms, ``Bag Space" and ``Embedded Space" \cite{amores13}. Bag Space methods define a distance or kernel function that determine the similarity between bags, while Embedded Space methods map bags into feature vectors which can then be used for classifiers \cite{amores13}. Embedded Space methods can be further divided into two subcategories: methods that simply aggregate information about all instances in a bag without differentiating them, and ``Vocabulary-based" methods that group certain similar instances together and then use those groups to form the feature vector \cite{amores13}. We use a vocabulary-based method in this paper, because having information about groups of similar sequence reads can be biologically important, as explained further in the clustering section.

@@

Some of the popular genome and metagenome assembly approaches include SOAPdenovo2 \cite{luo12}, IDBA-UD \cite{peng12}, Velvet \cite{zerbino08}, MetaVelvet \cite{namiki12}, ABySS \cite{simpson09}, and Ray Meta \cite{boisvert12}. We used SOAPdenovo2 in our study, because it was the assembler used in the original MGWAS study \cite{qin041012} and because it has been shown to be one of the fastest assembly algorithms \cite{peng12}. SOAPdenovo2 first constructs a type of directed graph called a \emph{de Bruijn} graph that represents the overlaps between different sequences \cite{li10}. Reads are divided into strings of length K called \emph{k-mers}; these k-mers are the nodes of the graph \cite{zerbino08}. The choice of K is up to the user, and is important for having good assembly results. The k-mer nodes in the graph have an edge between them if a read contains those k-mers in order with an overlap of K-1 nucleotides, and the direction of the edge indicates in which order the k-mers appear \cite{zerbino08}. SOAPdenovo2 then cleans this graph by removing nodes/sequences with few or no connections with other sequences, eliminating ``tips" that represent likely sequencing machine errors, and removing redundant edges \cite{li10}. This step helps to reduce the overall size of the data by eliminating some reads that would not have been useful anyway. The contigs are then formed by combining reads according to the de Bruijn graph: each contig represents a directed path in the graph \cite{zerbino08}. There are variations on how exactly de Bruijn graphs are represented and implemented, but Figure \ref{de-bruijn} illustrates the general process.

@@

UCLUST \cite{Edgar10}, CD-HIT \cite{Li01072006}, mothur \cite{schloss2009introducing}, DOTUR \cite{schloss2005introducing}, CROP \cite{Hao01032011}, MC-MinH \cite{sdm2013a}, and 
MC-LSH \cite{bibm2012} are some of the popular sequence clustering approaches. We use UCLUST within our study, which is one of the most widely used and cited metagenome clustering methods and has been shown to be amongst the most effective in terms of speed and accuracy in benchmarking studies \cite{bonder090112, sun042711}. 
%
